---
title: 'PS #3 Question 3'
author: 'David Zynda'
date: "February 5, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Previous Work

```{r, results='hide'}
## Problem 14

# (a) Generate Data 

x1 = rnorm(1000, 0, 1)
x2 = rnorm(1000, 0, 1)
e = rnorm(1000, 0, 1)

y = 1 + x1 + x2 + e

x1 = as.matrix(x1)
x2 = as.matrix(x2)
e = as.matrix(e)

y = as.matrix(y)


my_lm = lm(y ~ x1 + x2)
summary(my_lm)


# (b) compute residuals


t(my_lm$residuals) %*% x1
t(my_lm$residuals) %*% x2

## (c) verify true residuals are not orthoganol

t(e) %*% x1
t(e) %*% x2

## (d) Auxiliary Regression

aux = lm(x1 ~ x2)

solve(t(aux$residuals) %*% aux$residuals) %*% (t(aux$residuals) %*% y)


## What is E(x'x)

t(x1) %*% x1
```



## T-test

### Write a simulation program showing the t-test of the coefficient on $x_{1i}$ being 0 performs as it is derived. 

The simulation below will give an actual parameter value of $0$ for $x_1$ and $1$ for $x_2$. 

```{r}
x0 = rep(1, 1000)
x1 = rnorm(1000, 0, 1)
x2 = rnorm(1000, 0, 1)
e = rnorm(1000, 0, 1)

y = x0 + 0*x1 + x2 + e

X = as.matrix(cbind(x0, x1, x2))

beta = solve((t(X) %*% X)) %*% (t(X) %*% y)

y_hat = X %*% beta

res = y - y_hat
res_sq = res ** 2
stdev = sum(res_sq) / 998


## Display beta estimates: first row slope, second row estimate for beta1

beta


```




Compare this with the `lm()` package in `R`:

```{r}
summary(lm(y ~ x1 + x2))
```


Lastly, compute t-stat of $x_1$. The null hypothesis, of course, is that $\beta_1 = 0$. 

```{r}

std_error_beta = stdev * sqrt(1000 / (1000 * t(x1) %*% x1 - sum(x1)^2)) # page 12 http://www.stat.cmu.edu/~hseltman/309/Book/chapter9.pdf

t = (beta[2] - 0) / std_error_beta
t
```


The t statistic is ```r t```




## F-test


Generate Simulation

```{r}

x1 = rnorm(100, 0, 1)
x2 = rnorm(100, 0, 1)
e = rnorm(100, 0, 1)

y =  0 * x1 + 0 * x2 + e

```



```{r}

X = as.matrix(cbind(x1, x2))
beta = solve(t(X) %*% X) %*% t(X) %*% y

y_hat = X %*% beta 
res = y - y_hat
res2 = y - 0

sse_r = t(res2)%*%res2
sse_u = t(res)%*%res

Fstat = ( ( (sse_r - sse_u) / (100 - 98) ) /  (sse_u / 100))


```



The F-stat is ```r Fstat```. Hence, it is not signficant. 


__Significance at the 95% level:__
```{r}
qf(.95, df1=100, df2=98)
```











